

The code below reads the data set which includes up to the latest demand information (August 17) and prepares the data for the modeling process:

```{r}
library(fpp3)
library(tseries)
X <- read.csv("Booking Exercise.csv") %>%
  select(DATE, DEMAND, TUESDAY.BOOK, DOW.INDEX) %>%
  mutate(DATE = ymd(DATE)) %>%
  as_tsibble(index = DATE)
  


Fcst.X <- X %>% filter(DATE >= ymd("2010-08-18"))
X <- X %>% filter(DATE <= ymd("2010-08-17"))
```


1. Fit (a) an ETS, (b) a plain ARIMA, (c) a linear regression model to predict DEMAND as a function of TUESDAY.BOOK, (c) a non-seasonal regression model with ARIMA errors using TUESDAY.BOOK and DOW.INDEX as explanatory variables, and (d) a seasonal regression model with ARIMA errors using only TUESDAY.BOOK as an explanatory variable.


```{r}

X %>% autoplot(DEMAND)

X %>% ACF(DEMAND) %>%  autoplot() + labs(title = "DEMAND")
X %>% PACF(DEMAND) %>%  autoplot() + labs(title = "DEMAND")

X1 <- X %>%  mutate(diff.X = difference(DEMAND),diff2.X = difference(diff.X))

X1 %>% gg_tsdisplay(DEMAND, plot_type = "partial")  
X1 %>% gg_tsdisplay(diff.X, plot_type = "partial")

```


```{r}

X1 %>% features(DEMAND, unitroot_kpss) 
X1 %>% features(diff.X, unitroot_kpss) 
X1 %>% features(diff2.X, unitroot_kpss) 

X1 %>% features(DEMAND, unitroot_ndiffs)
X1 %>% features(DEMAND, unitroot_nsdiffs)


X1$DEMAND %>% adf.test()

X1$diff.X %>%
  na.omit() %>%  # Removing the leading NA
  adf.test()

X1$diff2.X %>%
  na.omit() %>%  # Removing the leading NA
  adf.test()

# Data is stationary, no differencing required

```

```{r}

m <- X %>% model(m_ETS = ETS(DEMAND),
                 m_ARIMA1 = ARIMA(DEMAND),
                 m_lr = TSLM(DEMAND ~ TUESDAY.BOOK),
                 m_ARIMA2 = ARIMA(DEMAND ~ TUESDAY.BOOK +	DOW.INDEX + pdq(1,0,0)),
                 m_ARIMA3 = ARIMA(DEMAND ~ TUESDAY.BOOK +	pdq(1,1,3) + PDQ(1,1,0)))

m %>% select(m_ETS) %>% report()
m %>% select(m_ARIMA1) %>% report()
m %>% select(m_lr) %>% report()
m %>% select(m_ARIMA2) %>% report()
m %>% select(m_ARIMA3) %>% report()

```


2. Which of the models are valid?

```{r}

m %>% augment() %>%
  features(.resid, ljung_box, lag = 10)

m %>% select(m_ETS) %>% gg_tsresiduals()
m %>% select(m_ARIMA1) %>% gg_tsresiduals()
m %>% select(m_lr) %>% gg_tsresiduals()
m %>% select(m_ARIMA2) %>% gg_tsresiduals()
m %>% select(m_ARIMA3) %>% gg_tsresiduals()

```
The Null Hypothesis of the Ljung-Box test is that residuals are uncorrelated. Hence, with a large p-value we can conclude that the residual independence cannot be rejected.

From the residual p-values, we can conclude that the ARIMA and ETS models have high p-value, but for the linear regression model we can conclude that the residuals are not independent. So there is a validity problem with the linear regression model.

```{r}

cor(X$DEMAND, X$TUESDAY.BOOK)
plot(DEMAND ~ TUESDAY.BOOK, data = X)


# TUESDAY.BOOK +	DOW.INDEX

```

3. Using the information criteria (AICc and BIC) what is the best model?

```{r}
m %>% glance() %>%
  select(.model, AIC, AICc, BIC)

```

As per IC, linear regression model is valid. Next one is a seasonal regression model with ARIMA errors using only TUESDAY.BOOK as an explanatory variable.

After looking at the correlation value, LJung Box data and the IC, we can select the seasonal regression model with ARIMA errors model.  

4. Use the Fcst.X data set created above to produce a demand forecast for the forthcoming week (i.e., Aug, 18 - 24) showing the default 80% and 90% confidence intervals.

```{r}

new_book_data <- new_data(X,7) %>%
  mutate(TUESDAY.BOOK = c(1657,1643,1124,1074,1839,1862,1886))

Forecast <- m %>% select(m_ARIMA3) %>%
  forecast(new_data = new_book_data)

Forecast %>% autoplot() + geom_line(X, mapping = aes(y = DEMAND))

```


5. What is your recommendation for your manager regarding the tour operator block?  Should your manager sell the block of 60 rooms at a discounted price?  Justify your answer.

```{r}

1877*0.67

# Date       Demand         Mean     Tuesday.book

# 2010-08-18 N(1775, 10332) 1775.397 1657

#2010-08-19 N(1740, 14673) 1739.562 1643

#2010-08-20 N(1162, 15206) 1161.549 1124

#2010-08-21 N(1170, 17355) 1169.913 1074

#2010-08-22 N(1644, 18206) 1643.513 1839

#2010-08-23 N(1549, 19959) 1549.144 1862

#2010-08-24 N(1816, 21015) 1816.182 1886

```

The cut-off for Aug 22 as per 67% service level is 0.67*1877 i.e., 1258 rooms. As per the forecast, in average the organic demand for rooms on Aug 22 (excluding the room block in question) is going to be 1643 which is already satisfying the minimum service level and can be sold at regular price. The condition is the block should be assigned to the tour operator only if the probability of organic demand is smaller than 67%. As the organic demand is greater than 67%, the manager should not sell the block of 60 rooms at a discounted price.


